Bird species composition in coastal urban habitats

### 2020-12-04 Ms_HabitatAves.Rmd
permanova

>>>>

```{r}
#Inv

Mod0 <- adonis2(AvesInv2 ~ 1, data = Amb, by = "terms")
Mod1 <- adonis2(AvesInv2 ~ Altitude +  Ocean +`Vegetation cover`, data = Amb, by = "terms")###################### Menor AIC
Mod2 <- adonis2(AvesInv2 ~ Altitude + `Paved surface`+ Shrublands+`Vegetation cover`, data = Amb, by = "terms")
Mod3 <- adonis2(AvesInv2 ~ Altitude + `Paved surface`+ Ocean , data = Amb, by = "terms")
Mod4 <- adonis2(AvesInv2 ~ Altitude +  Ocean +`Vegetation cover`+ Shrublands, data = Amb, by = "terms")
Mod5 <- adonis2(AvesInv2 ~ Altitude + `Paved surface`+ Shrublands+ Ocean, data = Amb, by = "terms")

AICc.PERMANOVA2 <- function(adonis2.model) {

  # check to see if object is an adonis2 model...

  if (is.na(adonis2.model$SumOfSqs[1]))
    stop("object not output of adonis2 {vegan} ")

  # Ok, now extract appropriate terms from the adonis model Calculating AICc
  # using residual sum of squares (RSS or SSE) since I don't think that adonis
  # returns something I can use as a likelihood function... maximum likelihood
  # and MSE estimates are the same when distribution is gaussian See e.g.
  # https://www.jessicayung.com/mse-as-maximum-likelihood/;
  # https://towardsdatascience.com/probability-concepts-explained-maximum-likelihood-estimation-c7b4342fdbb1
  # So using RSS or MSE estimates is fine as long as the residuals are
  # Gaussian https://robjhyndman.com/hyndsight/aic/ If models have different
  # conditional likelihoods then AIC is not valid. However, comparing models
  # with different error distributions is ok (above link).

  RSS <- adonis2.model$SumOfSqs[ length(adonis2.model$SumOfSqs) - 1 ]
  MSE <- RSS / adonis2.model$Df[ length(adonis2.model$Df) - 1 ]

  nn <- adonis2.model$Df[ length(adonis2.model$Df) ] + 1

  k <- nn - adonis2.model$Df[ length(adonis2.model$Df) - 1 ]

  # AIC : 2*k + n*ln(RSS/n)
  # AICc: AIC + [2k(k+1)]/(n-k-1)

  # based on https://en.wikipedia.org/wiki/Akaike_information_criterion;
  # https://www.statisticshowto.datasciencecentral.com/akaikes-information-criterion/ ;
  # https://www.researchgate.net/post/What_is_the_AIC_formula;
  # http://avesbiodiv.mncn.csic.es/estadistica/ejemploaic.pdf;
  # https://medium.com/better-programming/data-science-modeling-how-to-use-linear-regression-with-python-fdf6ca5481be

  AIC <- 2*k + nn*log(RSS/nn)
  AICc <- AIC + (2*k*(k + 1))/(nn - k - 1)

  output <- data.frame(AICc = AICc, k = k, N = nn)

  return(output)

}

DF<- data.frame(Modelo = paste0("Mod", 0:5), AICc = NA)

DF$AICc[1] <- AICc.PERMANOVA2(Mod0)$AICc
DF$AICc[2] <-  AICc.PERMANOVA2(Mod1)$AICc
DF$AICc[3] <- AICc.PERMANOVA2(Mod2)$AICc
DF$AICc[4] <-  AICc.PERMANOVA2(Mod3)$AICc
DF$AICc[5] <-  AICc.PERMANOVA2(Mod4)$AICc
DF$AICc[6] <-  AICc.PERMANOVA2(Mod5)$AICc

```{r}
#Prim

Mod0 <- adonis2(AvesPrim2 ~ 1, data = Amb, by = "terms")
Mod1 <- adonis2(AvesPrim2 ~ Altitude +  Ocean +`Vegetation cover`, data = Amb, by = "terms")
Mod2 <- adonis2(AvesPrim2 ~ Altitude + `Paved surface`+ Shrublands+`Vegetation cover`, data = Amb, by = "terms")
Mod3 <- adonis2(AvesPrim2 ~ Altitude + `Paved surface`+ Ocean , data = Amb, by = "terms")########################     Menor AIC
Mod4 <- adonis2(AvesPrim2 ~ Altitude +  Ocean +`Vegetation cover`+ Shrublands, data = Amb, by = "terms")
Mod5 <- adonis2(AvesPrim2 ~ Altitude + `Paved surface`+ Shrublands+ Ocean, data = Amb, by = "terms")

AICc.PERMANOVA2 <- function(adonis2.model) {

  # check to see if object is an adonis2 model...

  if (is.na(adonis2.model$SumOfSqs[1]))
    stop("object not output of adonis2 {vegan} ")

  # Ok, now extract appropriate terms from the adonis model Calculating AICc
  # using residual sum of squares (RSS or SSE) since I don't think that adonis
  # returns something I can use as a likelihood function... maximum likelihood
  # and MSE estimates are the same when distribution is gaussian See e.g.
  # https://www.jessicayung.com/mse-as-maximum-likelihood/;
  # https://towardsdatascience.com/probability-concepts-explained-maximum-likelihood-estimation-c7b4342fdbb1
  # So using RSS or MSE estimates is fine as long as the residuals are
  # Gaussian https://robjhyndman.com/hyndsight/aic/ If models have different
  # conditional likelihoods then AIC is not valid. However, comparing models
  # with different error distributions is ok (above link).

  RSS <- adonis2.model$SumOfSqs[ length(adonis2.model$SumOfSqs) - 1 ]
  MSE <- RSS / adonis2.model$Df[ length(adonis2.model$Df) - 1 ]

  nn <- adonis2.model$Df[ length(adonis2.model$Df) ] + 1

  k <- nn - adonis2.model$Df[ length(adonis2.model$Df) - 1 ]

  # AIC : 2*k + n*ln(RSS/n)
  # AICc: AIC + [2k(k+1)]/(n-k-1)

  # based on https://en.wikipedia.org/wiki/Akaike_information_criterion;
  # https://www.statisticshowto.datasciencecentral.com/akaikes-information-criterion/ ;
  # https://www.researchgate.net/post/What_is_the_AIC_formula;
  # http://avesbiodiv.mncn.csic.es/estadistica/ejemploaic.pdf;
  # https://medium.com/better-programming/data-science-modeling-how-to-use-linear-regression-with-python-fdf6ca5481be

  AIC <- 2*k + nn*log(RSS/nn)
  AICc <- AIC + (2*k*(k + 1))/(nn - k - 1)

  output <- data.frame(AICc = AICc, k = k, N = nn)

  return(output)

}

DF<- data.frame(Modelo = paste0("Mod", 0:5), AICc = NA)

DF$AICc[1] <- AICc.PERMANOVA2(Mod0)$AICc
DF$AICc[2] <-  AICc.PERMANOVA2(Mod1)$AICc
DF$AICc[3] <- AICc.PERMANOVA2(Mod2)$AICc
DF$AICc[4] <-  AICc.PERMANOVA2(Mod3)$AICc
DF$AICc[5] <-  AICc.PERMANOVA2(Mod4)$AICc
DF$AICc[6] <-  AICc.PERMANOVA2(Mod5)$AICc
